Review of Edge Detectors

### 2014-05-29 review/review_answer_2.tex

%\begin{abstract}
%\end{abstract}

\chapter{Introduction}
In this communication we answer to the second rounds of review of the article "A Contrario Selection of Optimal Partitions for Image Segmentation", submitted to the SIAM Journal on Imaging Sciences.
In the first place, we would like to thank again the reviewers for the detailed and thoughtful review provided. In the following we address the issues raised by the reviewers. For each issue, an answer is provided and a cross-reference with a red tag $C_n$ (where n is the number of the correction) to the corresponding (corrected) part of the paper is provided. In addition, changes made to the paper are highlighted in red (when significative) in the new version of the paper. 
These red marks will be removed for the final version.

The remarks from the reviewers are marked with the tag \textcolor{blue}{QUESTION} and our answers with the tag \ans.

%\section{Brief summary of changes}
\section{General changes}
\begin{itemize}
 \item Revised general writing of the article.
\item We added appendices to include secondary information that distracts from the main purpose of the article.
\item In that sense we moved the numeric values of the $\mathbf{mask_i}$ matrices to the first appendix
\item Removed several typos and orthographic errors.
\item improved code readme and demo.
\item  Changed the location of the demo, it is now located at: \url{http://dev.ipol.im/~juanc/ipol_demo/sc_edges/}
\end{itemize}

@@

\section{Paper review}

\que In general the paper provides all the information necessary. Some parts need nevertheless rewriting either because of minor mistakes (that make some sentences really hard to read) or because of mathematical inaccuracies.\\

\ans We thank the reviewer for the comments and for contributing to improve the quality of the paper. We did a major proof-reading of the paper and we did our best effort to improve the writing.\\

\que You mention a video but it was not included in the supplementary materials (35-757-1-SP.zip and 35-758-1-SP.zip) so that I could not see it.\\

@@

- compute the second order derivatives
- find the zero crossing locations and set them as edges

Once again the two algorithms differs mainly in the way of computing
derivatives. The former rely on the filtering with laplacian of gaussian or on
the filtering with a gaussian followed by a laplacian operator. The latter
relies on a local bicubic approximation followed by a second derivative
analytical computation. Sufficient details are again provided along with
algorithmic description. However some of the details could be made clearer (see
later comments).

The fourth section provides technical details on some filtering aspects and the
last section is dedicated to the experimental discussion.

\subsection{Detailed comments, typos and typesetting}
\subsubsection{Demo}
\que The url \url{http://dev.ipol.im/~haldos/ipol_demo/xxx_edges/} provided in the preprint
contains a broken demo. This is not very important though since the demo code
provided in the submission works correctly and it should eventually replace the
one specified in the pdf.\\

\ans The new demo is located  here \url{http://dev.ipol.im/~juanc/ipol_demo/sc_edges/}\\

### 2014-11-03 edge-detection-review-3rd-revision-v1.tex
First round of corrections for 3rd review. Changed almost every text change required. code still needs to be fixed

Edge detection is one of the oldest and most basic operations operations in image processing, which is often used as a basic building block for more complex algorithms.  for 
. The basic idea of edge detection is to detect abrupt changes in image intensity. 
Detecting those changes can be accomplished using first or second order derivatives. \\

@@

Algorithms based on first derivatives studied in this paper share a common structure, with the only difference  in the type of filtering used to compute those derivatives. Figure \ref{fig:blockdiagram1} shows a block diagram of that common structure. \myn{\label{correction_a:first_derivative:first_deriv_rephrase}\\

@@

\myn{\label{correction_a:first_derivative:taylor}}The most usual approach to approximate the first derivative is to use a the first order Taylor expansion with a small value of $h$. Thus $f'(x)$ is computed as
\begin{equation}
	f'(x) \simeq \frac{f(x+h) - f(x)}{h}.\\
\end{equation}
\myn{\label{correction_a:first_derivative:discretization}} Digital images are commonly seen as a discretization of a continuos image, where a pixel is evenly sample on a grid. If we denote the grid sizes by $\Delta x$ and $\Delta y$, the continuous coordinates $x$  and $y$ can be written as $x=i\Delta x$ and $y=i\Delta y$. When no information about the spatial sampling is available, the size of the grid it is commonly assumed to be $\Delta x=\Delta y=1$.\\
Thus, in the discrete space the first derivative of the image is computed as:
\begin{equation}
\label{eq:roberts1}
	f_x = \frac{\partial f(x,y)}{\partial x} \cong f(i+1,j) - f(i,j)
\end{equation}
and
\begin{equation}
\label{eq:roberts2}
	f_y = \frac{\partial f(x,y)}{\partial y} \cong f(i,j+1) - f(i,j).
\end{equation}

\myn{\label{correction_a:first_derivative:diagonal edges}} Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$ by filtering the image $f(x,y)$ with the 1-D masks shown in Figure \ref{fig:1dmasks}. However, in order to be able to compute diagonal edges, 2-D masks are needed. \\
The \textit{Roberts operators} are one of the earliest 
attempts to use 2-D masks for this purpose. These operators are based on computing the diagonal \tcom{differences} \myn{\label{correction_a:typo}} implemented by filtering an image with the masks in Figure \ref{fig:roberts}. By convention 
in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. It is also 
usual to scale the mask values, in order to have gain equal to $1$. \\

@@

\begin{SCfigure}[2][!h]
	\centering
	\begin{squarecells}{1}
		-1   \nline
		1  \nline
	\end{squarecells}
	\quad
	\begin{squarecells}{2}
		-1 & 1   \nline
	\end{squarecells}
	\caption{One-dimensional masks used to implement equations \ref{eq:roberts1} and \ref{eq:roberts2}.}
	\label{fig:1dmasks}
\end{SCfigure}

\begin{SCfigure}[1][!h]
	\centering
	\begin{squarecells}{2}
		-1 &  0  \nline
		0 & 1  \nline
	\end{squarecells}
	\quad
	\begin{squarecells}{2}
		0  & -1  \nline
		1 & 0  \nline
	\end{squarecells}
	\caption{\textit{Roberts cross-gradient} 2-D masks.}
	\label{fig:roberts}
\end{SCfigure}

@@

\begin{SCfigure}[1][!h]
	\centering
	\begin{squarecells}{3}
		$-\frac{1}{3}$ 	& $-\frac{1}{3}$ 	& $-\frac{1}{3}$	\nline
		0 			& 0			& 0			\nline
		$\frac{1}{3}$ 	& $\frac{1}{3}$ 	& $\frac{1}{3}$	\nline
	\end{squarecells}
	\quad
	\begin{squarecells}{3}
		$-\frac{1}{3}$ 	& 0 	& $\frac{1}{3}$	\nline
		$-\frac{1}{3}$	& 0	& $\frac{1}{3}$	\nline
		$-\frac{1}{3}$ 	& 0 	& $\frac{1}{3}$	\nline
	\end{squarecells}
	\caption{Normalized \textit{Prewitt} 2-D masks of size $3\times3$.}
	\label{fig:prewitt}
\end{SCfigure}

\myn{\label{correction_a:first_derivative:sobel}} The left mask can be constructed (up to a scale factor) as the convolution of two filters: a vertical derivative 
$[-1\ 0\ 1]$ and an horizontal moving average $[1\ 1\ 1]^T$. The same reasoning could be applied to the right mask. Hence, these operators can be interpreted as the consecutive application of an horizontal low pass filter with a vertical derivative filter. Thus, they show smoothing properties in the direction orthogonal to the gradient.

@@

\begin{SCfigure}[1][!h]
	\centering
	\begin{squarecells}{3}
		$-\frac{1}{4}$ 	& $-\frac{1}{2}$ 	& $-\frac{1}{4}$	\nline
		0 			& 0			& 0			\nline
		$\frac{1}{4}$ 	& $\frac{1}{2}$ 	& $\frac{1}{4}$	\nline
	\end{squarecells}
	\quad
	\begin{squarecells}{3}
		$-\frac{1}{4}$ 	& 0 	& $\frac{1}{4}$	\nline
		$-\frac{1}{2}$	& 0	& $\frac{1}{2}$	\nline
		$-\frac{1}{4}$ 	& 0 	& $\frac{1}{4}$	\nline
	\end{squarecells}
	\caption{\textit{Sobel} 2-D masks of size $3\times3$.}
	\label{fig:sobel}
\end{SCfigure}

@@

As mentioned before, edges can be seen as abrupt changes in intensity, i.e.\ higher values of gradient.
An example of this behavior is shown in Figure \ref{fig:profiles3d}, where gradient was computed using Sobel operators.

@@

Finally, the edges image is obtained by thresholding the gradient magnitude image, i.e.\ for pixel at position $i$ in the image compute
\begin{equation*}
	\mbox{edges}[i,j] = \begin{cases} 1,& \mbox{if}\ M[i,j]\geq\mbox{th} \\
									0,& \mbox{if}\ M[i,j]<\mbox{th}
					  \end{cases}
\end{equation*}
where $\mbox{th}$ is the threshold. A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}). 
It is noted that different thresholds lead to different results. Thicker or more edges for smaller thresholds, fewer edges 
for larger thresholds. This threshold, and the others mentioned below, are defined as a percentage of the maximum value 
of the gradient image. The advantage of this is to adapt the threshold to the dynamic range of the image. 
Therefore, the parameter $threshold$ in the algorithms take values ​​between $0$ and $1$.

@@

\begin{algorithm}[t!]
\caption{Marr-Hildreth edge detection algorithm.}
\label{algo:mh}
\begin{algorithmic}[1]
\REQUIRE input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $t_{ZC}$.
\STATE $im \leftarrow$ input image
\IF{Gaussian kernel}
	\STATE Define Laplacian operator $laplacian$
	\STATE $im_{LAPL} \leftarrow$ convolution($im$,$laplacian$)
\ELSE
	\STATE $kernel \leftarrow$ generate\_kernel($n$,$\sigma$) \COMMENT{Generated Gaussian or LoG kernel}\mynf{\label{correction_a:second_derivative:marr_algo}}.
	\STATE $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
	\STATE $im_{LAPL} \leftarrow im_{SMOOTHED}$
\ENDIF
\STATE $max_L \leftarrow 0$
\FORALL{pixel $i$ in image $im_{LAPL}$}
	\IF{$im_{LAPL}[i]>max_L$}
		\STATE $max_L \leftarrow im_{LAPL}[i]$
	\ENDIF
\ENDFOR
\FORALL{pixel $i$ in image $im_{LAPL}$, except borders}
	\FORALL{pair $(p_1,p_2)$ of opposite neighbors of $p$ in $im_{LAPL}$}
		\IF{($\mbox{sign}(im_{LAPL}[p_1])\neq\mbox{sign}(im_{LAPL}[p_2])$) \AND ($|im_{LAPL}[p_1]-im_{LAPL}[p_2]|>th_{ZC}$)}
			\STATE $im_{OUT}[i] \leftarrow 255$
		\ELSE
			\STATE $im_{OUT}[i] \leftarrow 0$
		\ENDIF
	\ENDFOR
\ENDFOR
\RETURN output image $\leftarrow im_{OUT}$
\end{algorithmic}
\end{algorithm}

@@

The surrounding neighborhood of a point $(x,y)$ in the image $f$ is approximated using the following bi-cubic polynomial
\begin{equation}
	\label{eq:bicubic}
	f(x,y) = k_1 + k_2x + k_3y + k_4x^2 + k_5xy + k_6y^2 + k_7x^3 + k_8x^2y + k_9xy^2 + k_{10}y^3. \\
\end{equation}
To solve this fitting problem, it is necessary to take more neighbors than coefficients to be adjusted. As there are 10 coefficients to compute, the smallest neighborhood of odd size that accomplishes this has size $5\times5$. Having $10$ coefficients and $25$ data 
points leads to an overdetermined system, which can be solved using least squares or any other fitting technique.\\

\myn{\label{second_derivative:haralick:statement}} In Haralick's work, this polynomial approximation is expressed in a local coordinate system centered on each pixel. In this way, the center of the neighboorhood is always $(0,0)$, and the coordinates $(x, y)$ are offsets from this center point (e.g. in a neighborhood of size 5x5, x and y take values between -2 and +2).\\
Then the solution is computed using least squares, and the local expression of the polynomial ensures that the values of the $k_i$ coefficientes are computed in the same way for every pixel in the image. Thus, they can be computed by convolving the image with a set of fixed and precomputed masks, which is a great speed-up of the algorithm. In the following we will state the approximation problem and show how to derive those masks from its solution.\\

Consider 25 points in a small neighborhood of a point $(x,y)$ in the image. This gives us an equal number of equations 
to find 10 coefficients. As mentioned before this leads to an \tcom{overdetermined} system which can be solved by least squares. 
By substituting the 25 data points into the polinomial equation (\ref{eq:bicubic}), the following system of 
equations is obtained,

@@

\myn{\label{second_derivative:haralick:normal}} Let us define a matrix $\mathbf{B} \in R^{10\times25}$ such that $\mathbf{B}=(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T$, thus

@@

\begin{equation}
	\label{eq:coefficients}
	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \hdots + b_{i,25}f_{25} \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{f} . \mathbf{b_i},\\
\end{equation}

where $b_i$ is the vector containing the $i-th$ row of $B$, i.e.\myn{\label{correction_a:second_derivative:haralick_bi}}

@@

Let us rearrange $b_i$ as a 5x5 matrix and denote it by $\mathbf{mask_i}$ as follows
\begin{equation}
	\label{eq:b_i}
	\mathbf{mask_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \hdots	& b_{i,5}	\\
									b_{i,6}		& b_{i,7}	& \hdots	& b_{i,10}	\\
									\vdots		& \vdots	& \ddots	& \vdots	\\
									b_{i,21}	 & b_{i,22}	& \hdots	& b_{i,25}	\\
					\end{bmatrix}.\\
\end{equation}

\vspace{1ex}
\myn{\label{second_derivative:haralick:mask_computation}} We will compute one value of $k_i$ for each pixel on the image. Thus we can store these coefficientes in a matrix $\bf{K_i}$ of the same size as the original image. Then, using the mask in shown in Equation \ref{eq:b_i} we can obtain $\bf{K_i}$ as the convolution of the original image with $\mathbf{mask_i}$. \\
\begin{equation}
	\label{eq:convolution_b}
	\mathbf{K_i} = f \star \mathbf{mask_i} \: \mathtt{for} \: i \in \{1 \ldots 10\}\\
\end{equation}

@@

\myn{\label{second_derivative:haralick:analitical:condition}} The main idea of Haralick's algorithm, as stated in the original work, is to find "negative sloped zero crossings of the second derivative". Thus, and pixel will lie on an edge if the following three conditions hold\footnote{Please note that there is a typo in Haralick's original paper. The statement of the idea is presented at the beggining of Section III, in the first column of page 42. The three conditions are presented at the end of the same section, in the second column. There it reads $f''_{\alpha}(\rho)<0$, $f''_{\alpha}(\rho)=0$ and $f''_{\alpha}(\rho)<0$. Clearly, the first and second conditions can't be true at the same time, thus the first one should be $f'''_{\alpha}(\rho)<0$.}:

@@

\myn{\label{second_derivative:haralick:analitical:statement}} After the neighborhood of each point of the image is approximated using the bi-cubic polynomial expression in equation \ref{eq:bicubic}, the idea is to compute the derivatives of the image in the direction of the gradient using that approximation.\\

Let $\theta$ be the gradient angle, defined with respect clockwise from the y-axis\footnote{Although it seems a non standard choice, we followed exactly from the original work}. We \tcom{approximate} it using the first order terms of the polynomial \ref{eq:bicubic} obtaining the following expression:

@@

\myn{\label{second_derivative:haralick:analitical:derivative}} Let us consider the line $r$ passing by $(x,y)$ with angle $\theta$. The coordinates of each point $(u,v)$ on this line can be described in terms of  the angle $\theta$ and the distance $\rho$ to $(x,y)$ as follows

\begin{equation*}
	x = \rho\cos{\theta} \: \text{and} \: y = \rho\sin{\theta} \\
\end{equation*}

in the bi-cubic polynomial, we obtain

@@

\myn{}The simplest way to impose these conditions is starting with the third one, i.e. that the third derivative has to be negative. Thus,\

@@

By construction $\theta$ is the direction of the gradient, which means that the edge will always be an ascending step. Thus by definition the function $f_{\theta}(\rho)$ is non-decreasing on $\rho=0$. $f'_{\theta}(0) \geq 0$. This implies that $C_1 \geq 0$.

@@

If we evaluate the solution $\rhoˆ*$ on it, we obtain

\begin{equation}\label{eq:cond:first}
f'_{\theta}(\rho) = C_1 - \frac{C_2^2}{C_3}  \\
\end{equation}

@@

In Haralick's work $\rho_0$ is a fixed threshold which acts as an argument to the algorithm.
\myn{\label{second_derivative:haralick:analitical:parameter_value}}  This parameter must be greater than zero and and less than the size of the grid (i.e. $\rho_0<1$). In this work we obtained the best results with values between $0.4$ and $0.6$.\\

@@

\begin{algorithm}[t]
\caption{Haralick edge detection algorithm.}
\label{algo:haralick}
\begin{algorithmic}[1]
\REQUIRE input image (width $w$, height $h$), edge point condition threshold $\rho_0$.BibTex | RIS | RefWorks
\STATE $im \leftarrow$ input image
\STATE Define the ten $5x5$ masks ($mask_1\dots mask_{10}$) used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See Table \ref{table:b_i}.}
\FORALL{pixel $i$ in image $im$}
	\STATE $neighbors \leftarrow$ $5\times5$ neighborhood of pixel $i$ in image $im$
	\FOR{$j=1$ \TO $10$}
		\STATE $k_j \leftarrow$ convolution($neighbors$,$mask[j]$)
	\ENDFOR
	\STATE $C_2 \leftarrow \frac{k_2^2k_4 + k_2k_3k_5 + k_3^2k_6}{k_2^2 + k_3^2}$
	\STATE $C_3 \leftarrow \frac{k_2^3k_7 + k_2^2k_3k_8 + k_2k_3^2k_9 + k_3^3k_{10}}{(\sqrt{k_2^2 + k_3^2})^3}$
	\IF{{$|C_2/3C_3|\leq\rho_0$} \AND {$C_3<0$}}
		\STATE $im_{OUT}[i] \leftarrow 255$
	\ELSE
		\STATE $im_{OUT}[i] \leftarrow 0$
	\ENDIF
\ENDFOR
\RETURN output image $\leftarrow im_{OUT}$
\end{algorithmic}
\end{algorithm}

@@

\myn{\label{}} All implemented algorithms use some common mathematical operations that are independent of the algorithms themselves. 
These operations, although basic, have a great impact on the algorithm outcome, and thus they need to be 
implemented with care. In this section those operations are reviewed.

@@

The size of the kernel $n$ and the standard deviation of the exponential function $\sigma$ are both 
input parameters, but these are not strictly independent of each other. 
A value of $n$ large enough is needed to ensure that no information is lost when creating the kernel $G$. To ensure this, $n$ is taken equal to the first odd integer greater than $6\sigma$. Larger values of $n$ do not add more significant samples ​​of $G$, and increase the number of operations in the convolution.

>>>>

Edge detection is one of the oldest and most basic operations operations in image processing, which is often used as a basic building block for more complex algorithms. The basic idea of edge detection is to detect abrupt changes in image intensity. 
Detecting those changes can be accomplished using first or second order derivatives. \\

@@

Algorithms based on first derivatives studied in this paper share a common structure, with the only difference  in the type of filtering used to compute those derivatives. Figure \ref{fig:blockdiagram1} shows a block diagram of that common structure. \\

@@

The most usual approach to approximate the first derivative is to use a the first order Taylor expansion with a small value of $h$. Thus $f'(x)$ is computed as
\begin{equation}
	f'(x) \simeq \frac{f(x+h) - f(x)}{h}.\\
\end{equation}
Digital images are commonly seen as a discretization of a continuous image, where a pixel is evenly sampled on a grid. If we denote the grid sizes by $\Delta x$ and $\Delta y$, the continuous coordinates $x$  and $y$ can be written as $x=i\Delta x$ and $y=i\Delta y$. When no information about the spatial sampling is available, the size of the grid it is commonly assumed to be $\Delta x=\Delta y=1$.\\
Thus, in the discrete space the first derivative of the image is computed as:
\begin{equation}
\label{eq:roberts1}
	f_x = \frac{\partial f(x,y)}{\partial x} \cong f(i+1,j) - f(i,j)
\end{equation}
and
\begin{equation}
\label{eq:roberts2}
	f_y = \frac{\partial f(x,y)}{\partial y} \cong f(i,j+1) - f(i,j).
\end{equation}

Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$ by filtering the image $f(x,y)$ with the 1-D masks shown in Figure \ref{fig:1dmasks}. However, in order to be able to compute diagonal edges, 2-D masks are needed. \\
The \textit{Roberts operators} are one of the earliest 
attempts to use 2-D masks for this purpose. These operators are based on computing the diagonal differences implemented by filtering an image with the masks in Figure \ref{fig:roberts}. By convention 
in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. It is also 
usual to scale the mask values, in order to have gain equal to $1$. \\

@@

\begin{figure}[2][!h]
	\centering
	\begin{squarecells}{1}
		-1   \nline
		1  \nline
	\end{squarecells}
	\quad
	\begin{squarecells}{2}
		-1 & 1   \nline
	\end{squarecells}
	\caption{One-dimensional masks used to implement equations \ref{eq:roberts1} and \ref{eq:roberts2}.}
	\label{fig:1dmasks}
\end{figure}

\begin{figure}[1][!h]
	\centering
	\begin{squarecells}{2}
		-1 &  0  \nline
		0 & 1  \nline
	\end{squarecells}
	\quad
	\begin{squarecells}{2}
		0  & -1  \nline
		1 & 0  \nline
	\end{squarecells}
	\caption{\textit{Roberts cross-gradient} 2-D masks.}
	\label{fig:roberts}
\end{figure}

@@

\begin{figure}[1][!h]
	\centering
	\begin{squarecells}{3}
		$-\frac{1}{3}$ 	& $-\frac{1}{3}$ 	& $-\frac{1}{3}$	\nline
		0 			& 0			& 0			\nline
		$\frac{1}{3}$ 	& $\frac{1}{3}$ 	& $\frac{1}{3}$	\nline
	\end{squarecells}
	\quad
	\begin{squarecells}{3}
		$-\frac{1}{3}$ 	& 0 	& $\frac{1}{3}$	\nline
		$-\frac{1}{3}$	& 0	& $\frac{1}{3}$	\nline
		$-\frac{1}{3}$ 	& 0 	& $\frac{1}{3}$	\nline
	\end{squarecells}
	\caption{Normalized \textit{Prewitt} 2-D masks of size $3\times3$.}
	\label{fig:prewitt}
\end{figure}

The left mask can be constructed (up to a scale factor) as the convolution of two filters: a vertical derivative 
$[-1\ 0\ 1]$ and an horizontal moving average $[1\ 1\ 1]^T$. The same reasoning could be applied to the right mask. Hence, these operators can be interpreted as the consecutive application of an horizontal low pass filter with a vertical derivative filter. Thus, they show smoothing properties in the direction orthogonal to the gradient.

@@

\begin{figure}[1][!h]
	\centering
	\begin{squarecells}{3}
		$-\frac{1}{4}$ 	& $-\frac{1}{2}$ 	& $-\frac{1}{4}$	\nline
		0 			& 0			& 0			\nline
		$\frac{1}{4}$ 	& $\frac{1}{2}$ 	& $\frac{1}{4}$	\nline
	\end{squarecells}
	\quad
	\begin{squarecells}{3}
		$-\frac{1}{4}$ 	& 0 	& $\frac{1}{4}$	\nline
		$-\frac{1}{2}$	& 0	& $\frac{1}{2}$	\nline
		$-\frac{1}{4}$ 	& 0 	& $\frac{1}{4}$	\nline
	\end{squarecells}
	\caption{\textit{Sobel} 2-D masks of size $3\times3$.}
	\label{fig:sobel}
\end{figure}

@@

As mentioned before, edges can be seen as abrupt changes in intensity, i.e. higher values of gradient.
An example of this behavior is shown in Figure \ref{fig:profiles3d}, where gradient was computed using Sobel operators.

@@

Finally, the edges image is obtained by thresholding the gradient magnitude image, i.e. for pixel a at position $i,j$ in the image compute
\begin{equation*}
	\mbox{edges}[i,j] = \begin{cases} 1,& \mbox{if}\ M[i,j]\geq\mbox{th} \\
									0,& \mbox{if}\ M[i,j]<\mbox{th}
					  \end{cases}
\end{equation*}
where $\mbox{th}$ is the threshold. A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}). 
It is noted that different thresholds lead to different results. Thicker or more edges for smaller thresholds, fewer edges 
for larger thresholds. This threshold, and the others mentioned below, are defined as a percentage of the maximum value 
of the gradient image. The advantage of this is to adapt the threshold to the dynamic range of the image. 
Therefore, the parameter $threshold$ in the algorithms takes values ​​between $0$ and $1$.

@@

\begin{algorithm}[t!]
\caption{Marr-Hildreth edge detection algorithm.}
\label{algo:mh}
\begin{algorithmic}[1]
\REQUIRE input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $t_{ZC}$.
\STATE $im \leftarrow$ input image
\IF{Gaussian kernel}
	\STATE Define Laplacian operator $laplacian$
	\STATE $im_{LAPL} \leftarrow$ convolution($im$,$laplacian$)
\ELSE
	\STATE $kernel \leftarrow$ generate\_kernel($n$,$\sigma$) \COMMENT{Generated Gaussian or LoG kernel}.
	\STATE $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
	\STATE $im_{LAPL} \leftarrow im_{SMOOTHED}$
\ENDIF
\STATE $max_L \leftarrow 0$
\FORALL{pixel $i$ in image $im_{LAPL}$}
	\IF{$im_{LAPL}[i]>max_L$}
		\STATE $max_L \leftarrow im_{LAPL}[i]$
	\ENDIF
\ENDFOR
\FORALL{pixel $i$ in image $im_{LAPL}$, except borders}
	\FORALL{pair $(p_1,p_2)$ of opposite neighbors of $p$ in $im_{LAPL}$}
		\IF{($\mbox{sign}(im_{LAPL}[p_1])\neq\mbox{sign}(im_{LAPL}[p_2])$) \AND ($|im_{LAPL}[p_1]-im_{LAPL}[p_2]|>th_{ZC}$)}
			\STATE $im_{OUT}[i] \leftarrow 255$
		\ELSE
			\STATE $im_{OUT}[i] \leftarrow 0$
		\ENDIF
	\ENDFOR
\ENDFOR
\RETURN output image $\leftarrow im_{OUT}$
\end{algorithmic}
\end{algorithm}

@@

The surrounding neighborhood of a point $(x_0,y_0)$ in the image $f$ is approximated using the following bi-cubic polynomial
\begin{align}
	\label{eq:bicubic:long}
	f(x,y) = k_1 + k_2 (x-x_0) + k_3(y-y_0) + k_4(x-x_0)^2 + k_5(x-x_0)(y-y_0)  +\\
	 k_6(y-y_0) ^2 + k_7(x-x_0)^3 + k_8(x-x_0)^2(y-y_0)  + k_9(x-x_0)(y-y_0) ^2 + k_{10}(y-y_0) ^3 \\
\end{align}
To solve this fitting problem, it is necessary to take more neighbors than coefficients to be adjusted. As there are 10 coefficients to compute, the smallest neighborhood of odd size that accomplishes this has size $5\times5$. Having $10$ coefficients and $25$ data 
points leads to an overdetermined system, which can be solved using least squares or any other fitting technique.\\

In Haralick's work, this polynomial approximation is expressed in a local coordinate system centered on each pixel. In this way, the center of the neighboorhood $(x_0,y_0)$ is always $(0,0)$, and the coordinates $(x, y)$ are offsets from this center point (e.g. in a neighborhood of size $5\times5$, x and y take values between -2 and +2). Thus, the expression of the polynomial is\\

\begin{equation}
	\label{eq:bicubic}
	f(x,y) = k_1 + k_2x + k_3y + k_4x^2 + k_5xy + k_6y^2 + k_7x^3 + k_8x^2y + k_9xy^2 + k_{10}y^3. \\
\end{equation}

Then the solution is computed using least squares, and the local expression of the polynomial ensures that the values of the $k_i$ coefficients are computed in the same way for every pixel in the image. Thus, they can be computed by convolving the image with a set of fixed and precomputed masks, which results in a great speed-up of the algorithm. In the following we will state the approximation problem and show how to derive those masks from its solution.\\

Consider 25 points in a small neighborhood the point $(0,0)$. This gives us an equal number of equations to find 10 coefficients. As mentioned before this leads to an overdetermined system which can be solved by least squares. 
By substituting the 25 data points into the polynomial equation (\ref{eq:bicubic}), the following system of 
equations is obtained,

@@

Let us define a matrix $\mathbf{B} \in R^{10\times25}$ such that $\mathbf{B}=(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T$, thus

@@

\begin{equation}
	\label{eq:coefficients}
	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \hdots + b_{i,25}f_{25} \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{b_i}\mathbf{f} ,\\
\end{equation}

where $b_i$ is the vector containing the $i-th$ row of $B$, i.e.

@@

Let us rearrange $b_i$ as a $5\times5$ matrix and denote it by $\mathbf{mask_i}$ as follows
\begin{equation}
	\label{eq:b_i}
	\mathbf{mask_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \hdots	& b_{i,5}	\\
									b_{i,6}		& b_{i,7}	& \hdots	& b_{i,10}	\\
									\vdots		& \vdots	& \ddots	& \vdots	\\
									b_{i,21}	 & b_{i,22}	& \hdots	& b_{i,25}	\\
					\end{bmatrix}.\\
\end{equation}

\vspace{1ex}
We will compute one value of $k_i$ for each pixel on the image. Thus we can store these coefficients in a matrix $\bf{K_i}$ of the same size as the original image. Then, using the mask shown in Equation \ref{eq:b_i} we can obtain $\bf{K_i}$ as the convolution of the original image with $\mathbf{mask_i}$. \\
\begin{equation}
	\label{eq:convolution_b}
	\mathbf{K_i} = f \star \mathbf{mask_i} \: \mathtt{for} \: i \in \{1 \ldots 10\}\\
\end{equation}

@@

The main idea of Haralick's algorithm, as stated in the original work, is to find "negative sloped zero crossings of the second derivative". This means that it is actually looking for ascending step edges. According to Haralick's definition, a pixel will lie on such an edge if the following three conditions hold\footnote{Please note that there is a typo in Haralick's original paper. The statement of the idea is presented at the beggining of Section III, in the first column of page 42. The three conditions are presented at the end of the same section, in the second column. There it reads $f''_{\alpha}(\rho)<0$, $f''_{\alpha}(\rho)=0$ and $f'_{\alpha}(\rho)\neq 0$. Clearly, the first and second conditions can't be true at the same time, thus the first one should be $f'''_{\alpha}(\rho)<0$.}:

@@

After the neighborhood of each point of the image is approximated using the bi-cubic polynomial expression in equation \ref{eq:bicubic}, the idea is to compute the derivatives of the image in the direction of the gradient using that approximation.\\

Let $\theta$ be the gradient angle, defined clockwise with respect to the y-axis\footnote{Although it seems a non standard choice, we followed exactly from the original work}. We approximate it using the first order terms of the polynomial \ref{eq:bicubic} obtaining the following expression:

@@

Let us consider the line $r$ passing by $(0,0)$ with angle $\theta$. The coordinates of each point $(x,y)$ on this line can be described in terms of  the angle $\theta$ and the distance $\rho$ to $(0,0)$ as follows

\begin{align*}
	x = \rho\cos{\theta} \\
	y = \rho\sin{\theta} 
\end{align*}

If we substitute the above expressions in the bi-cubic polynomia of eq., we obtain

@@

The simplest way to impose the above-mentioned conditions is starting with the third one, i.e. that the third derivative has to be negative. Thus,

@@

By construction $\theta$ is the direction of the gradient, which means that the edge will always be an ascending step. Thus by definition the function $f_{\theta}(\rho)$ is non-decreasing in a neighborhood of $\rho=0$, which means that $f'_{\theta}(0) \geq 0$. This in turn implies that $C_1 \geq 0$.

@@

If we evaluate $f'_{\theta}$ on the solution $\rhoˆ*$ using the above formula, we obtain

\begin{equation}\label{eq:cond:first}
f'_{\theta}(\rhoˆ*) = C_1 - \frac{C_2^2}{C_3}  \\
\end{equation}

@@

In Haralick's work $\rho_0$ is a fixed threshold which acts as an argument to the algorithm.
This parameter must be greater than zero and and less than the size of the grid (i.e. $\rho_0<1$). In this work we obtained the best results with values between $0.4$ and $0.6$.\\

@@

\begin{algorithm}[t]
\caption{Haralick edge detection algorithm.}
\label{algo:haralick}
\begin{algorithmic}[1]
\REQUIRE input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
\STATE $im \leftarrow$ input image
\STATE Define the ten $5\times5$ masks ($mask_1\dots mask_{10}$) used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See Table \ref{table:b_i}.}
\FORALL{pixel $i$ in image $im$}
	\STATE $neighbors \leftarrow$ $5\times5$ neighborhood of pixel $i$ in image $im$
	\FOR{$j=1$ \TO $10$}
		\STATE $k_j \leftarrow$ convolution($neighbors$,$mask[j]$)
	\ENDFOR
	\STATE $C_2 \leftarrow \frac{k_2^2k_4 + k_2k_3k_5 + k_3^2k_6}{k_2^2 + k_3^2}$
	\STATE $C_3 \leftarrow \frac{k_2^3k_7 + k_2^2k_3k_8 + k_2k_3^2k_9 + k_3^3k_{10}}{(\sqrt{k_2^2 + k_3^2})^3}$
	\IF{{$|C_2/3C_3|\leq\rho_0$} \AND {$C_3<0$}}
		\STATE $im_{OUT}[i] \leftarrow 255$
	\ELSE
		\STATE $im_{OUT}[i] \leftarrow 0$
	\ENDIF
\ENDFOR
\RETURN output image $\leftarrow im_{OUT}$
\end{algorithmic}
\end{algorithm}

@@

All implemented algorithms use some common mathematical operations that are independent of the algorithms themselves. 
These operations, although basic, have a great impact on the algorithm outcome, and thus they need to be 
implemented with care. In this section those operations are reviewed.

@@

The size of the kernel $n$ and the standard deviation of the exponential function $\sigma$ are both 
input parameters, but these are not strictly independent of each other. 
The value of $n$ needs to be large enough to ensure that no information is lost when creating the kernel $G$. To ensure this, $n$ is taken equal to the first odd integer greater than $6\sigma$. Larger values of $n$ do not add more significant samples ​​of $G$, and increase the number of operations in the convolution.
